---
title: "Portfolio_Outline"
author: "Olivia Townsend"
date: "11/1/2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Files
I have a combination of files to analyze. One file is an Excel version of my field notes, containing information like nest number and band number, deploy time/date and recovery time/date of GPS loggers, and morphometric measurements (mass, culmen length, head and bill length, wing chord) for each tagged bird. I also have a .csv file of tracking data for each bird, which includes latitude, longitude, speed, and altitude for every sampling point. I tracked 22 birds total.

## Importing the Files

```{r import_fieldnotes}
library(xlsx)
setwd("~/Desktop/RTTR 2017")
rttr<-read.xlsx("RTTR_2017.xlsx",sheetIndex = 1)
```

```{r import_trackingdata}
fnames = dir("~/Desktop/RTTR 2017/csv files", pattern = ".csv",full.names = TRUE)
for (i in 1:length(fnames)){
  temp=read.csv(fnames[i])
  myfile=basename(fnames[i])
  loc=regexpr(pattern="[[:digit:]]{2}_",text = myfile)
  birdID = substr(myfile,start=loc[[1]],stop = loc[[1]]+1)
  temp$BIRDID= birdID
  if (i==1){
    rttrall = temp
  } else {
    rttrall = rbind(rttrall,temp)}
}
```


## Final Products

# Morphometrics
In addition to creating a table of the morphometric data, I am creating a plot in order to display trends between different measurements (e.g. mass vs. head and bill length) and the sexes. In tropicbirds, males are typically larger than females and I would like to see if my sampled birds show the same pattern. For now, I am using a 3D scatterplot to graph mass vs. head and bill length vs. wing chord. Once I receive the molecular sexing results, I can color the data points by sex in order to differentiate between males and females.
```{r morphometrics}
library(xlsx)
library(scatterplot3d)
library(knitr)
library("kableExtra")

setwd("~/Desktop/RTTR 2017")
rttr<-read.xlsx("RTTR_2017.xlsx",sheetIndex = 1)

#table
kable(rttr[,c(1,13,15,16)], col.names = c("BIRD ID", "MASS (g)", "HEAD AND BILL (mm)", "WING CHORD (mm)"), align = "c")

#graph
scatterplot3d(x=rttr$HEAD...BILL..mm.,y=rttr$WING.CH..mm.,z=rttr$MASS..g.,
              xlab = "Head and Bill Length (mm)",ylab = "Wing Chord (mm)",
              zlab = "Mass (g)",cex.symbols=1,cex.axis=0.8,pch=19,y.margin.add = 0.1)

kable(rttr[,c(1,13,15,16)], col.names = c("BIRD ID", "MASS (g)", "HEAD AND BILL (mm)", "WING CHORD (mm)"), align = "c")

```

# Map of all foraging tracks
```{r all_tracks}
library(maps)
library(mapdata)
library(lubridate)
library(dplyr)
library(argosfilter)
library(RColorBrewer)
data("world2MapEnv")
data("worldMapEnv")

map(database="world",xlim=c(-162,-150),ylim=c(16,26),mar = c(3.5,1,1,1),fill = TRUE,col = "darkgray")
grid(nx=NULL,ny=NULL,col = "darkgray",lty = 1)
axis(1, at=c(-162,-160,-158,-156,-154,-152,-150),cex.axis=0.8,labels = c(expression(162*degree*W),
      expression(160*degree*W),expression(158*degree*W),
      expression(156*degree*W),expression(154*degree*W),
      expression(152*degree*W),expression(150*degree*W)))
axis(2, at=c(16,18,20,22,24,26),cex.axis=0.8,labels = c(expression(16*degree*N),
      expression(18*degree*N),expression(20*degree*N),expression(22*degree*N),
      expression(24*degree*N),expression(26*degree*N)))
lines(rttrall$Longitude,rttrall$Latitude,col="#8856A7")
points(-157.68,21.28,col="#FECC5C",pch=17)
```

# Map of tracks from one individual showing differences in foraging strategy (short trips and long trip)
```{r individual_track}
map(database="world",xlim=c(-159,-156),ylim=c(17.5,22),mar = c(3.5,1,1,1),fill = TRUE,col = "darkgray")
grid(nx=NULL,ny=NULL,col = "darkgray",lty = 1)
axis(1, at=c(-159,-158,-157,-156),cex.axis=0.8,labels = c(expression(159*degree*W),
      expression(158*degree*W),expression(157*degree*W),
      expression(156*degree*W)))
axis(2, at=c(17,18,19,20,21,22),cex.axis=0.8,labels = c(expression(17*degree*N),
      expression(18*degree*N),expression(19*degree*N),expression(20*degree*N),
      expression(21*degree*N),expression(22*degree*N)))
lines(rttr04$Longitude,rttr04$Latitude,col="#8856A7")
points(-157.68,21.28,col="#FECC5C",pch=17)
```

# Trip Parameters
Similar to the morphometric data, I would rather put this information into a figure rather than a table. To characterize foraging behavior, I would like to calculate mean trip duration, distance, and max range of short and long foraging trips. I also want to compare parameters between the sexes. I do not have the code for this section yet, but I know I want my output figure to be a bar graph with SD bars included.
```{r trip_parameters, eval=FALSE}
library(adehabitatLT)
library(oce)
library(geosphere)


```

# Classification of Behavioral States
I would like to generate a map that shows behavioral states (rest, transit, area-restricted search) for each track. I do not have the code yet but know I will need to use the Residence in Space and Time (RST) algorithm to classify the behaviors. I may also plot daily activity patterns (stacked histogram?) and bar graphs of mean speed and altitude by behavioral state and time of day.
```{r behavioral_states, eval=FALSE}
library(adehabitatLT)
library(lubridate)
library(dplyr)
library(argosfilter)
library(ggplot2)
library(maptools)

```

# Spatial Distribution
To visualize areas used for foraging, I plan to create a kernel density utilization distribution map (25-95% UDs), or perhaps one for short trips and one for long trips. Again, no code for this yet.
```{r kernelUD, eval=FALSE}
library(adehabitatHR)
library(ggplot2)
library(maps)
library(mapdata)
library(dplyr)

```

# Stable Isotopes
Lastly, I want to plot stable carbon and nitrogen values of blood and feather samples (and possibly distinguished by sex) to look at diet characteristics between the breeding and non-breeding season. However, this depends on when I get my stable isotope results back from the lab. No code for this yet but if I get my results back in time, it will be a basic scatterplot with carbon isotope values on the x-axis, nitrogen isotope values on the y-axis, and SD included on each plot point.
```{r stable_isotopes}

```

