---
title: "Week_7_Homework"
author: "Olivia Townsend"
date: "10/4/2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import the CSU salary database and convert the "Employee.Name" column to a character class

```{r import and convert}
df = read.csv("../data/california-state-university-2015.csv")
df$Employee.Name = as.character(df$Employee.Name)
```

# Find the entries...

## with the name "Martinez"
```{r Martinez}
my_pattern = "[M][a][r][t][i][n][e][z]"
df$Employee.Name[grepl(pattern = my_pattern, df$Employee.Name)]
```

## whose name starts with the letters "Wil"*
```{r name starts with Wil}
my_pattern = "[W][i][l]"
df$Employee.Name[grepl(pattern = my_pattern, df$Employee.Name)]
```

## with a hyphenated name
```{r hyphenated name}
my_pattern = "[[:lower:]][-][[:upper:]]"
df$Employee.Name[grepl(pattern = my_pattern, df$Employee.Name)]
```

## whose middle name is Paul
```{r middle name Paul}
my_pattern = "[[:blank:]][P][a][u][l][[:blank:]]"
df$Employee.Name[grepl(pattern = my_pattern, df$Employee.Name)]
```

## with the last name Martinez
```{r last name with Martinez}
my_pattern = "[[:blank:]][M][a][r][t][i][n][e][z]"
df$Employee.Name[grepl(pattern = my_pattern, df$Employee.Name)]
```

## with the last name Martinez but no hyphenated names or Jr.
```{r last name just Martinez}
my_pattern = "[[:blank:]][M][a][r][t][i][n][e][z]$"
df$Employee.Name[grepl(pattern = my_pattern, df$Employee.Name)]
```

## with first name that starts with "E", has a middle initial, and last name Martinez
```{r E middle initial Marinez}
my_pattern = "^[E]\\w*[[:blank:]][A-Z][[:blank:]][M][a][r][t][i][n][e][z]$"
df$Employee.Name[grepl(pattern = my_pattern, df$Employee.Name)]
```

## with first name that starts with "E", has NO middle initial, and last name Martinez
```{r E no middle initial Martinez}
my_pattern = "^[E]\\w*[[:blank:]][M][a][r][t][i][n][e][z]$"
df$Employee.Name[grepl(pattern = my_pattern, df$Employee.Name)]
```
